---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(maps)
library(ggfortify)
```

```{r read-clean-oecd-data}
oecd_violence <- 
  read_csv("violence.csv") %>% 
  filter(TIME=="2019") %>% 
  select(LOCATION, SUBJECT, Value) %>% 
  rename(location=LOCATION, indicator=SUBJECT, value=Value) 
```


```{r join-regions-data}
##merging data so i have the continents/regions of the world
regions_data<-
  read_csv("regionsdata.csv") %>% 
  select(-"...6", -"SpecialNotes")


regions_data <- rename(regions_data, c("location" = "Country Code"))

#Merging datasets

violence_with_region <- merge(oecd_violence, regions_data, by="location")

# Change indicators to their own columns
violence_region_wide <- violence_with_region %>% 
  pivot_wider(names_from="indicator", values_from="value") %>% 
  mutate(LAWDOMVIOL = as.character(LAWDOMVIOL)) %>% 
  mutate(LAWDOMVIOL = fct_relevel(LAWDOMVIOL, "0.25", "0.5",  "0.75", "1")) 

# Reordering group factor levels for IncomeGroup
violence_region_wide$IncomeGroup <- factor(violence_region_wide$IncomeGroup,
                         levels = c("High income", "Upper middle income", "Lower middle income", "Low income"))

#LAWDOMVIOL2019 <- violence_with_region %>% 
 # filter(SUBJECT=="LAWDOMVIOL", TIME==2019) %>% 
 # mutate(LAWDOMVIOL=Value)

#ATTITUDEVIOL2019 <- violence_with_region %>% 
 # filter(SUBJECT=="ATTITUDEVIOL", TIME==2019) %>% 
 # mutate(ATTITUDEVIOL=Value)

#PREVVIOLLIFETIME2019 <- violence_with_region %>% 
# filter(SUBJECT=="PREVVIOLLIFETIME", TIME==2019) %>% 
#  mutate(PREVVIOLLIFETIME=Value)

#mergeddata1 <- merge(LAWDOMVIOL2019, ATTITUDEVIOL2019, by="LOCATION")
#finalmergeddata<- merge(mergeddata1,PREVVIOLLIFETIME2019, by="LOCATION" )
```

```{r}

#violence_region_wide<- finalmergeddata[c(1,6,9,10,14,27,40)]

##Organized the Dataset in a way in which countries are in the same row

#Exploring LAWDOMVIOL by region and incomegroup
ggplot(violence_region_wide, aes(x=LAWDOMVIOL, fill=IncomeGroup))+
  geom_bar()+
  facet_wrap(~IncomeGroup)+
  theme(legend.position = "none")+
  labs (
    title="Laws on Domestic Violence by Country Income Group",
    subtitle="where the closer to 1, the more laws or practices discriminate against women’s 
rights",
    x="Laws on Domestic Violence Indicator",
    y="Count"
  )
  
## all regression models should run in the same dataset! pay attention to NA's
```
#Analysis of Grap 1: Laws on Domestic Violence by Country Income Group
#The graph shows that most countries, regardless of income group, fall in the "0.75" measure for Laws on Domestic Violence. That means that their laws and practices are unsuficcient to garantee the well-being of women. It is evident that even though there are few countries in which the number reaches the value 1 (laws and practices are completely discriminatory against women), the realization that most countries in the global still have unsuficcient laws that protect women's rights is incredibly worrying. We can also observe that there are more countries with better laws and practices for women's rights among high income countries (that is, countries in which the indicator equals 0.25). This also holds true in a lesser extent for upper middle income countries. For both lower middle income and low income, however, there seems to be more countries in the 0.5 and 0.75 levels then in the 0.25. This unvels a pattern that establishes that higher income coutries, despite also having large number of countries that have a indicator of law on domestic violence equal to 0.75, have also higher number of countries that have a somewhat benefitial laws and practices against domestic violence. However, this does not seem to be a causal effect. It is highly unlikely that a countrie's income group alone establishes the laws on domestic violence, but it points us in the direction of establishing a connection between higher levels of economic develpment and better legislation on domestic violence.
```{r}
ggplot(violence_region_wide)+
  geom_bar(mapping=aes(x=LAWDOMVIOL, fill=Region))+
  facet_wrap(~Region, ncol=3)+
  theme(legend.position = "none")+
  labs (
    title="Laws on Domestic Violence by Region of the World",
    subtitle="where the closer to 1, the more laws or practices discriminate against women’s 
rights",
    x="Laws on Domestic Violence Indicator",
    y="Count"
  )
?filter
filter(violence_region_wide, IncomeGroup=="High income")
#48 high income countries
filter(violence_region_wide, IncomeGroup=="Upper middle income")
#43 upper middle income
filter(violence_region_wide, IncomeGroup=="Low income")
#25 Low income countries
filter(violence_region_wide, IncomeGroup=="Lower middle income")
#45 lower middle income countries
```

The problem with doing this split by regions is that the number of countries per category fluctuates a lot. There are only two high income countries compared to many Sub-Saharan African countries, which limit our possibilities of seeing patterns. 

```{r}
#Exploring ATTITUDEVIOL by region and incomegroup
ggplot(violence_region_wide)+
  geom_histogram(mapping=aes(x=ATTITUDEVIOL, fill=IncomeGroup))+
  facet_wrap(~IncomeGroup)+
  theme(legend.position = "none")+
  labs (
    title="Attitude toward violence by Income Group",
    subtitle="The percentage of women who agree that a husband/partner is justified in 
beating his wife/partner under certain circumstances",
    x="Share of Women (%)",
    y="Count"
  )
```
There seems to be a clear association between attitude towards violence against women and a countrie's group income. High and upper middle income countries seem to have a lower share of women who agree that the husband is justified in inflicted violence in their wives. What are the cultural, social, and political factors that seem to change that attitude in lower middle and low income countries?
```{r}
ggplot(violence_region_wide)+
  geom_histogram(mapping=aes(x=ATTITUDEVIOL, fill=Region))+
  facet_wrap(~Region)+
  theme(legend.position = "none")+
  labs (
    title="Attitude toward violence by Region",
    subtitle="The percentage of women who agree that a husband/partner is justified in 
beating his wife/partner under certain circumstances",
    x="Share of Women (%)",
    y="Count"
  )
```

```{r}
#Exploring PREVVLIFETIME by region and incomegroup
ggplot(violence_region_wide)+
  geom_histogram(mapping=aes(x=PREVVIOLLIFETIME, fill=IncomeGroup))+
  facet_wrap(~IncomeGroup)+
  theme(legend.position = "none")+
  labs (
    title="Prevalence of violence in the lifetime by IncomeGroup",
    subtitle="The percentage of women who have experienced physical and/or sexual 
violence from an intimate partner at some time in their life",
    x="Share of Women (%)",
    y="Count"
  )
  
```

```{r}
ggplot(violence_region_wide)+
  geom_histogram(mapping=aes(x=PREVVIOLLIFETIME, fill=Region), bins=40)+
  facet_wrap(~Region)+
  theme(legend.position = "none")+
  labs (
    title="Prevalence of violence in the lifetime by Region",
    subtitle="The percentage of women who have experienced physical and/or sexual 
violence from an intimate partner at some time in their life",
    x="Share of Women (%)",
    y="Count"
  )
```

```{r}
#Weakness: not a lot of NA and MENA region countries

#Plotting data 

ggplot(violence_region_wide, aes(x = LAWDOMVIOL, y = ATTITUDEVIOL, color=Region)) +
  geom_point(alpha=0.5)+
  labs(title = "Relationship between Law on domestic violence and attitudes towards violence", x = "Law on domestic violence", y = "attitudes towards violence")
```

```{r}
#barstack comparing this data by incomegroup

ggplot(violence_region_wide, 
       aes(x = LAWDOMVIOL, 
           fill = IncomeGroup)) + 
  geom_bar(position = "fill")+
labs (
   title="Laws on Domestic Violence by Country Income Group",
    subtitle="where the closer to 1, the more laws or practices discriminate against 
women’s rights",
    x="Laws on Domestic Violence Indicator",
    y="Count"
  )

#ggplot(violence_region_wide, 
#       aes(x = LAWDOMVIOL, 
#           fill = IncomeGroup)) + 
# geom_bar(position = "dodge")

#ggplot(violence_region_wide, 
#       aes(x = LAWDOMVIOL, 
#           fill = IncomeGroup)) + 
 # geom_bar(position = "stack")
```

```{r}
ggplot(violence_region_wide, 
       aes(x = ATTITUDEVIOL, 
           fill = IncomeGroup,
           )) + 
  geom_histogram(alpha=0.8, position = "stack", bins=50)+
 labs(
   title="Prevalence of violence in the lifetime by Region",
    subtitle="The percentage of women who have experienced physical and/or sexual 
violence from an intimate partner at some time in their life",
    x="Share of Women (%)",
    y="Count"
)
  
```

```{r}
ggplot(violence_region_wide, 
       aes(x = PREVVIOLLIFETIME, 
           fill = IncomeGroup,
           )) + 
  geom_histogram(alpha=0.8, position = "stack", bins=50)+
  labs (
    title="Prevalence of violence in the lifetime by income group",
    subtitle="The percentage of women who have experienced physical and/or sexual 
violence from an intimate partner at some time in their life",
    x="Share of Women (%)",
    y="Count"
  )
```

```{r}
#barstack by region 

##this is not very useful
#ggplot(violence_region_wide, 
#       aes(x = LAWDOMVIOL, 
 #          fill = Region)) + 
#  geom_bar(position = "dodge")

#ggplot(violence_region_wide, 
     #  aes(x = PREVVIOLLIFETIME, 
      #     fill = Region)) + 
#  geom_histogram(position = "stack")

#ggplot(violence_region_wide, 
#       aes(x = ATTITUDEVIOL, 
 #          fill = Region)) + 
 # geom_bar(position = "stack")

```

```{r}
#Plotting data 

ggplot(violence_region_wide, aes(x = LAWDOMVIOL, y = ATTITUDEVIOL)) +
  geom_point()+
  labs(title = "Relationship between Law on domestic violence and attitudes towards violence", x = "Law on domestic violence", y = "attitudes towards violence")

###

ggplot(violence_region_wide, aes(x = LAWDOMVIOL, y = PREVVIOLLIFETIME)) +
  geom_point()+
  labs(title = "Relationship between Law on domestic violence and Prevalance in Lifetime", x = "Law on domestic violence", y = "Prevalance in lifetime")

###

ggplot(violence_region_wide, aes(x = PREVVIOLLIFETIME, y = ATTITUDEVIOL)) +
  geom_point()+
  labs(title = "Relationship between Prevelance on lifteime and attitudes towards violence", x = "Prevalance on lifetime", y = "attitudes towards violence")
```

```{r}
#Problems and limits with the data
#- causation and correlation: Even though we find that NA and European countries ( high income countries) have better laws on domestic violence, as well as less worse attitudes on violence against women, and lower prevalence of women who experience abuse, this does not really tell us that these countries are "better" for women or men 
#- questions that indicators are asking versus what they are trying to measure, analyse each indicator and how they are measure the stuff they are measuring
#ex: prevalence in lifetime: they are only calculating violence coming from intimate partner, what about the levels of harassment that come from outside intimate circles?
#ex: law on domestic violence, how do you quantify that?
#ex: attitude towards violence: what questions are they asking. "the percentage of women who agree that a husband/partner is justified in beating his wife/partner under certain circumstances" why is this only women's complicity on it? is this really the best question to ask around attitude towards violence?

```


```{r read-country-coordinates}
# Read country coordinates
country_coordinates <- read_csv("countries_codes_and_coordinates.csv")

# Clean country coordinates
country_coordinates <- country_coordinates %>% 
  select(Country, `Alpha-3 code`, `Latitude (average)`, `Longitude (average)`) %>%
  rename(country=Country, alpha_code=`Alpha-3 code`, lat=`Latitude (average)`, long=`Longitude (average)`) %>% 
  distinct(alpha_code, .keep_all= TRUE)

# Join country coordinates with violence data to get longitude and latitude
violence_region_coordinates <- 
  violence_with_region %>% 
  left_join(country_coordinates, by= c("location" = "alpha_code"))

violence_region_coordinates_wide <- 
  violence_region_wide %>% 
  left_join(country_coordinates, by= c("location" = "alpha_code"))
```


```{r}
violence_region_coordinates %>% 
  filter(indicator!= "LAWDOMVIOL") %>% 
  ggplot(aes(x=value, fill=indicator)) + 
  geom_histogram(alpha=0.8)
```


```{r compare-all-three-indicators}
violence_region_coordinates_wide %>%
  ggplot(aes(x=ATTITUDEVIOL, y=PREVVIOLLIFETIME, color=LAWDOMVIOL)) +
  geom_point(size=5, alpha=0.8) +
  labs(
    title = "Prevalence of Violence Against Attitudes Towards Violence",
    subtitle = "Grouped by Laws on Domestic Violence",
    x="Attitude Towards Violence",
    y="Prevalence of Violence in Lifetime",
    color="Laws on Dom Violence"
) +
  theme_gray() +
  scale_colour_viridis_d(direction=-1)

```

```{r join-un-data}

# Read UN data on prevalence of violence in past 12 months
un_women_violence <- read_csv("un_women_violence_data.csv") %>% 
  select(-`Source Link`) %>% 
  rename(country = `Country Name`, un_year = `Reference year`, un_prev_viol_12m = `48. Proportion of women subjected to physical and/or sexual violence by a current or former intimate partner, in the last 12 months`)

un_oecd_women_violence_wide <- violence_region_coordinates_wide %>% 
  left_join(un_women_violence, by=c("country"="country"))
```

```{r pivot-un-oecd-data-longer}
# Pivot longer
un_oecd_women_prev_violence <- un_oecd_women_violence_wide %>%
  select(-LAWDOMVIOL, -ATTITUDEVIOL)  %>% 
  rename(un_prev_viol = un_prev_viol_12m, oecd_prev_viol = PREVVIOLLIFETIME) %>% 
  pivot_longer(c("oecd_prev_viol", "un_prev_viol"), names_to="indicator", values_to="prev_viol")
```


```{r comparison-map, layout="l-body-outset", fig.height=7}
# Compare OECD data with UN data
world_map <- map('world', fill=TRUE, plot=FALSE)

labels <- list(
  "oecd_prev_viol"="OECD Prevalence of Violence (lifetime)",
  "un_prev_viol" ="UN Prevalence of Violence (12 months)"
)

facet_labeller <- function(variable,value){
  return(labels[value])
}

autoplot(world_map, geom = 'polygon', col="white", fill="grey", ylim=c(-54,85), xlim=c(-160, 200)) +
  geom_point(data=un_oecd_women_prev_violence, aes(x=long, y=lat, color=prev_viol, size=prev_viol), alpha=0.8) +
  theme_minimal() +
  scale_colour_viridis_c(direction=-1) +
  labs(
    size="",
    color = "",
    title="Comparison Between OECD and UN on Prevalence of Domestic Violence Against Women",
  ) +
  facet_wrap(~indicator,
             labeller = facet_labeller,
             ncol=1)
```



```{r}
# Compare OECD data with UN data
un_oecd_women_violence_wide %>% 
  ggplot(mapping=aes(x=PREVVIOLLIFETIME, y=un_prev_viol_12m, color=LAWDOMVIOL)) +
  geom_point(alpha=0.8, size=5) +
  scale_colour_viridis_d(direction=-1) +
  labs(
    x="Prevalence of Violence (OECD)",
    y="Prevalence of Violence (UN)",
    color="Laws on Domestic Violence",
    title="UN vs OECD Prevalence of Domestic Violence"
  )
```

```{r corr-un-oecd}
# Compare OECD data with UN data
cor(un_oecd_women_violence_wide$PREVVIOLLIFETIME, un_oecd_women_violence_wide$un_prev_viol_12m, use="complete.obs")
```

# Introduction

# Theory/Hypotheses

# Data

# Analysis and Interpretation of Results

# Conclusions





